Historical trends in the contribution of six major causes of death to the black-white

### 2017-04-13 Manuscript/Manuscript_COD_Trends.Rmd
words

---
title: Historical trends in the contribution of six major causes of death to the black-white
  life expectancy gap by US state
bibliography: bibliography.bib
output:
  word_document: default
  html_document: default
always_allow_html: yes
---

@@

@@

To describe the general pattern in these trend lines we used random effects meta regression. Each state's mean contribution was regressed against a linear spline for calendar time that was estimated using four internal knots. The model included random effects for state and was inversely-weighted by the standard error of each state's mean estimate within each year. To identify divergent states, we compared the state-specific trends (estimated separately for each state using the Bayesian model) to the general pattern estimated using meta regression.

*Need to add information on the trends in mortality* -- where?

@@

Injury is one of the least common causes of death. Over time, injury-related mortality has decreased precipitously among blacks, although an [insurgence][better word?] of crime resulted in heightened mortality from injuries among black males during the 1990s (**Figure X**). For whites, most states exhibit an initial period of gradual decreases in injury mortality, followed by a period of gradual increases, especially among white females (**Figure X**). Thus, injury mortality rates among blacks are the same as, or lower than those among whites in many states, with white females demonstrating sharper increases. Thus, the contribution of injuries to the mortality gap has generally decreased over time, and has become negative in those states with lower injury-related mortality among blacks than whites (**Figures 9 and 10**). Among males, DC, Illinois, Michigan and Missouri exhibit the largest recent injury-related contributions. Among females, many southern states exhibit negative contributions (indicating lower injury-related mortality among black women), while DC, Illinois, and Michigan having higher contributions than the general trend.

@@

- In this paper, we examined state-level trends in six major CODs' contributions to the black-white LE gap.
- While we found that many state-level trends followed the same pattern of COD-specific contributions, several states were divergent. Furthermore, different patterns emerged for males compared to females.
-

>>>>

---
title: Historical trends in the contribution of six major causes of death to the black-white
  life expectancy gap by US state
bibliography: bibliography.bib
output:
  html_document: default
  word_document: default
always_allow_html: yes
---

@@

We also calculated age-standardized mortality rates for year-state-race-sex stratum, to aid the interpretation of the decomposition results. We used the US standard population from the year 2000 [ref].

@@

We used random effects meta regression to describe the general pattern in these trend lines. Each state's mean contribution was regressed against a linear spline for calendar time that was estimated using four internal knots. The model included random effects for state and was inversely-weighted by the standard error of each state's mean estimate within each year. To identify divergent states, we compared the state-specific trends (estimated separately for each state using the Bayesian model) to the general pattern estimated using meta regression.

@@

Injury is one of the least common causes of death. Over time, injury-related mortality has decreased precipitously among blacks, although an [insurgence][better word?] of crime resulted in heightened mortality from injuries among black males during the 1990s (**eFigure X**). For whites, most states exhibit an initial period of gradual decreases in injury mortality, followed by a period of gradual increases, especially among white females (**eFigure X**). Thus, injury mortality rates among blacks are the same as, or lower than those among whites in many states, with white females demonstrating sharper increases. Thus, the contribution of injuries to the mortality gap has generally decreased over time, and has become negative in those states with lower injury-related mortality among blacks than whites (**Figures 9 and 10**). Among males, DC, Illinois, Michigan and Missouri exhibit the largest recent injury-related contributions. Among females, many southern states exhibit negative contributions (indicating lower injury-related mortality among black women), while DC, Illinois, and Michigan having higher contributions than the general trend.

@@

- In this paper, we examined state-level trends in six major CODs' contributions to the black-white LE gap.

- While we found that many state-level trends followed the same pattern of COD-specific contributions, several states were divergent. For every causes of death except communicable disease, Massachusettes and New York exhibited steep declines to the contribution of each cause to the life expectancy gap, implying that decreasing inequality across a broad range of causes has results in eliminating the life expectancy gap in those states.

- On the other hand, DC underwent an increasing life expectancy gap over time and has the largest life expectancy difference of all the states. This analysis
